Theories of deliberated choice

### 2023-11-23 simple.tex
More about indicativeness

A claim $C$ is \emph{informative} iff $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.
A claim $C$ is \emph{informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) ≠ P \cup \set{\emptyset}$.
A claim $C$ is \emph{strongly informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) \ssub P$.
% A claim $C$ is \emph{indicative} iff $\exists H \in \finite(\incr(\allhists)), i \in I \suchthat \pi(\set{{\allleadstog} \suchthat H \wsub \ileadstoinvg(P)} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
A claim $C$ is \emph{indicative} iff $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub \suchthat \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
\begin{remark}
  Informativeness permits mere counting: the claim that exactly two persons deliberately believe $\phi$ is informative.
  Informativeness may be a desirable property (rather than informativeness beyond counting) as one may want to know whether a given proposition is believed by a majority of individuals, or whether a given proposition is believed by nobody.
  Note also that strong informativeness beyond counting implies informativeness beyond counting and indicativeness.
  Note also that informativeness beyond counting implies informativeness.
  \Cref{th:infoBeyonCKO} (and extensions, \cref{th:infoBeyonCKOext,th:strinfoBeyonCKO}) shows that informativeness beyond counting (and variants) is too demanding if one wants to investigate situations where individuals have differing deliberate preferences.
\end{remark}

@@

\begin{theorem}
  Assume that $I$ is finite and $P_{i_1} = \phi$ and $\forall i \in I \setminus \set{i_1}: P_i = ¬\phi$.

  Define $a_1$ as an argument decisive in favor of $\phi$ for $i_1$ and $a_i$ as arguments decisive in favor of $¬\phi$ for $i \in I \setminus \set{i_1}$.

  Assume (wlog) that $I = \set{i_1, …, i_{\card{I}}}$.
  Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{I}}$, $\hist_i = (\hist_{i - 1}, a_1, a_i)$.
  We know that, $\forall i \in \intvl{2, \card{I}}$: $\hist_i \ileadsto[i_1] \phi$ and $\hist_i \ileadsto[i] ¬\phi$.
  Define $\theta = \set{(\hist_i, i_1, \phi) \suchthat i \in \intvl{2, \card{I}}}$.

  Finally, the attitudes in $C$ compatible with $\theta$ are only $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C = \set{{\allleadstog} \suchthat {\ileadstog[i_1]} = {\ileadsto[i_1]}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} = {\ileadsto[i_1]}})(i_1) = \phi$.
\end{proof}

@@

  Finally, the attitudes in $C$ compatible with $\theta$ are only $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C \wsub \set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat j \in I_1}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat j \in I_1}})(i_1) = \phi$.
\end{proof}

>>>>

A claim $C$ is \emph{informative} iff $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.
A claim $C$ is \emph{informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) ≠ P \cup \set{\emptyset}$.
A claim $C$ is \emph{strongly informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) \ssub P$.
% A claim $C$ is \emph{indicative} iff $\exists H \in \finite(\incr(\allhists)), i \in I \suchthat \pi(\set{{\allleadstog} \suchthat H \wsub \ileadstoinvg(P)} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
A claim $C$ is \emph{indicative} iff $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub \suchthat \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
\begin{remark}
  Informativeness permits mere counting: the claim that exactly two persons deliberately believe $\phi$ is informative.
  Informativeness may be a desirable property (rather than informativeness beyond counting) as one may want to know whether a given proposition is believed by a majority of individuals, or whether a given proposition is believed by nobody.
  Note also that strong informativeness beyond counting implies informativeness beyond counting and indicativeness.
  Note also that informativeness beyond counting implies informativeness.
  \Cref{th:infoBeyonCKO} (and extensions, \cref{th:infoBeyonCKOext,th:strinfoBeyonCKO}) shows that informativeness beyond counting (and variants) is too demanding if one wants to investigate situations where individuals have differing deliberate preferences.
\end{remark}
\begin{remark}
  We could weaken indicativeness by allowing for disjuncts: observing either that $\hist \ileadsto \phi$ or that $\hist \ileadsto \psi$ permits to deduce something about $P_i$.
  (Or more complex formulas under the form of a set of possible functions in $P^H$?)
  (Actually I am not sure that this really weakens the condition.)
\end{remark}

@@

@@

  Finally, the attitudes in $C$ compatible with $\theta$ are $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C \wsub \set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat j \in I_1}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat j \in I_1}})(i_1) = \phi$.
\end{proof}

\begin{theorem}
  \label{th:suffIndic}
  Define $I_1 = \set{i \in I \suchthat P_i = \phi}$ and $I_2 = \set{i \in I \suchthat P_i = ¬\phi}$. Assume that $I = I_1 \cup I_2$.
  % Assume that for some $H \in \finite(\incr(\allhists)): [\exists i_1 \in I_1 \suchthat \forall \hist \in H: \hist \ileadsto[i_1] \phi] \land [\forall i \in I_2: \exists \hist \in H \suchthat \hist \ileadsto ¬\phi]$.
  Assume that for some $H \in \finite(\incr(\allhists))$ and $i_1 \in I_1$, we have $\forall i \in I_2$ that $\exists \hist \in H \suchthat {\allleadsto}(\hist, i_1) \neq {\allleadsto}(\hist, i_2)$.
  Then, there exists a claim that holds and is anonymous and indicative.
  \commentOC{In fact $I_2$ must include all those whose dj is not $\phi$.}
\end{theorem}
\begin{proof}
  Define the claim as $\forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}$
  and let us show that the claim is indicative for $i = i_1$ and $\theta = \set{(\hist, i_1, {\allleadsto}(\hist, i_1)) \suchthat \hist \in H}$.

  Finally, the attitudes in $C$ compatible with $\theta$ are $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C \wsub \set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat j \in I_1}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat j \in I_1}})(i_1) = \phi$.
\end{proof}
\begin{corollary}
  Define $I_1 = \set{i \in I \suchthat P_i = \phi}$ and $I_2 = \set{i \in I \suchthat P_i = ¬\phi}$. Assume that $I = I_1 \cup I_2$.
  If $\exists A \in \finite(\allargs) \suchthat \forall i \in I_2, \exists a \in A \suchthat a \ileadstost ¬\phi$, then there exists a claim that holds and is anonymous and indicative.
\end{corollary}
\begin{proof}
  We know that $i_1$ has some decisive argument $a_1$ in favor of $\phi$.
  Assume (wlog) that $A = \set{2, …, \card{A} + 1}$.
  Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{A} + 1}$, $\hist_i = (\hist_{i - 1}, a_1, i)$.
  Define $H = \set{\hist_i \suchthat i \in \intvl{2, \card{A} + 1}}$.
  Apply \cref{th:suffIndic}.
\end{proof}
